The markings of the trade: exploring the patterns of olive oil production in Roman Baetica

### 2020-04-30 Paper/jaccardbaetica.tex
first draft

%cambiar quizás
\begin{abstract}

@@

%primer párrafo más general---- especifico (¿dónde se ha aplicado?)
\xavi{queda raro esto de "en arqueología la cultura material nos permite ver la distribución en el Mediterráneo". Yo iría directa a hablar del Mediterráneo sin la primera frase.}

%spread of goods and ideas had a enormous impact during the Roman Empire. The progressive exploitation of communication networks allowed a major interaction between communities
%This can be seen by the fact that material culture is found in different regions along the Mediterranean as a frequent indicator of trade exchange in the archaeological record.

@@

%Material culture is one of the most frequent indicators of trade in the archaeological record. In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}. Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}\xavi{aquí faltan citas de obras más generales; esto es un ejemplo pero muy específico no?}.

@@

\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}

@@

We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province. The stamp database was compiled by CEIPAC group \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}). However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information. Consequently, we discard integrate the fragmented stamps in our dataset. We finally filter a total sample of 987 stamps composed of 130 different stamps from 81 workshops.

@@

Morisita-Horn index was used to compute a matrix to group sites with the stamps codes that share similarities. The matrix was then visualised using hierarchical clustering. The algorithm was selected to cluster similar groups in order to analyse the relationship between groups of sites and the distribution of similar stamp codes.

\maria{mejorar}

@@

%crear como una introducción

>>>>

\begin{abstract}

@@

%\xavi{queda raro esto de "en arqueología la cultura material nos permite ver la distribución en el Mediterráneo". Yo iría directa a hablar del Mediterráneo sin la primera frase.}

@@

%Material culture is one of the most frequent indicators of trade in the archaeological record. In archaeology, they allow us to highlight a part of the mechanism of production and distribution of goods along the Mediterranean \citep{bevan_mediterranean_2014}. Particularly, the spread of these factors had an important impact during the Roman Age, when the progressive exploitation of communication networks allowed a major interaction between communities \citep{orengo_seeds_2016}\xavi{aquí faltan citas de obras más generales; esto es un ejemplo pero muy específico no?}. 
%spread of goods and ideas had a enormous impact during the Roman Empire. The progressive exploitation of communication networks allowed a major interaction between communities
%This can be seen by the fact that material culture is found in different regions along the Mediterranean as a frequent indicator of trade exchange in the archaeological record.

@@

%\xavi{creo que le falta chicha aquí, sobre qué se ha hecho hasta ahora, qué quieres saber específicamente, el caso de estudio...}

@@

We studied a dataset of 3798 stamps collected from different Dressel 20 amphora workshops in \textit{Baetica} province. The stamp database was compiled by CEIPAC group \citep{remesal_centro_2015} (see CEIPAC database here \url{http://romanopendata.eu}). However, approximately the 70 \% of stamps cannot be tested due to fragmentation or incomplete information. Consequently, we discard integrate the fragmented stamps in our dataset. We finally filtered a total sample of 987 stamps composed of 130 different stamps from 81 workshops.

@@

Morisita-Horn index resulted in a numeric matrix of dissimilarities containing the distance value between the sites and the code stamps. The method was used to group sites with similar stamps codes. The matrix was then computed using hierarchical clustering. The algorithm was selected to cluster similar groups in order to analyse the relationship between groups of sites and the distribution of similar stamp codes. The results were visualised using a dendrogram to detect groups of sites sharing similar stamp codes.

@@
